{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-09-25 thesis/overview.tex
start wirting glossary

This section provides a convenient reference of terminology used throughout the thesis.

\begin{enumerate}
	\item
\end{enumerate}

% concurrency bug
% race condition: "a confusing term which in some circles means concurrency bug and others data race, which we will avoid"
% data race
% atomicity violation (what even is this.)
% false positive false negative
% thread
% transition
% conflict
% independent: "see conflict"; and note diff b/w this (applies to transitions) equivalent (applies to interleavings)
% state space
% interleaving
% DPOR
% estimation (WBE, RE)
% ETA
% landslide
% quicksand
% iterative deepening
% MC, stateless MC, systematic concurrency testing (TODO: stop calling it SMC start saying SCT)
% TA: teaching assistant
% P2 (note abotu psu - dosent call it p2, but we call it that for both schools here
% happens before, all 3 versions - DPOR, LHB, and PHB (TODO: replace PHB uses of 'happens before' with vector clock
% userspace, kernelspace
% "landslide friendly"

>>>>

This section provides a convenient reference of terminology used throughout the thesis.
Use of {\em italics} within definitions will indicate terms also defined in this list.

\begin{enumerate}
	\item {\bf API.}
		Short for {\em application programming interface} (which nobody ever says).
		% TODO
	\item {\bf Atomic.}
		Said of a sequence of operations within a single thread,
		whose intermediate state cannot be observed by or interfered with
		by certain conflicting operations from another thread.
		Must always be qualified as ``atomic {\em with respect to} X'',
		where X is the set of possibly-conflicting operations to be protected against.
		Apart from hardware-provided atomic assembly instructions (e.g., {\tt xchg}),
		i.e., for any multi-instruction sequence,
		setting X as ``everything'' can be surprisingly difficult to achieve safely in practice;
		for example, disabling interrupts in kernel code does not prevent other CPUs' simultaneous execution.
	\item {\bf Atomicity violation.}
		A term introduced by AVIO \cite{avio} to refer to a particular class of {\em concurrency bug}
		in which the programmer's expectation of {\em atomicity} around a certain sequence of operations
		is violated by another thread interleaving in between.
		This is distinct from language-level {\em data races},
		which may result in bugs by being reordered by the compiler or hardware;
		% TODO: put a section reference for weak memory
		% mb have to write a section on it here
		however, this thesis checks programs at the executable level and assumes sequential consistency,
		so all data races that still result in bugs (e.g., unprotected {\tt x++})
		must technically also be atomicity violations.
		Also, this thesis is focused on identifying failures, not root causes,
		so will simply avoid the term entirely.
	\item {\bf Branch.}
		Another term for {\em interleaving}, with an emphasis on the {\em execution tree}
		interpretation of {\em state spaces}.
		May also refer to an entire subtree outside of the current interleaving,
		characterized by the choice to run a certain {\em thread} at a certain past {\em preemption point},
		comprising all possible interleavings which may arise with that as a prefix.
	\item {\bf Concurrency bug.}
		A program behaviour which manifests nondeterministically depending on the thread scheduling pattern,
		which is not desired or intended by the programmer.
		This thesis further assumes these to be identifiable by computable predicates on the execution state,
		excluding some real bugs which may result in data corruption that never actually leads to crashes or assertion
		failures under the test case in question.
	\item {\bf Conflict.}
		% TODO
	\item {\bf Data race.}
		An unsynchronized pair of memory accesses between two threads that can, may, or did execute concurrently.
		{\em Race} for short.
		Distinguished further in this thesis as {\em limited happens-before} races and {\em pure happens-before} races;
		see \sect{\ref{sec:background-datarace}}.
		Distinct from the more nebulously-defined {\em concurrency bug},
		in that they are identified precisely by analysis on execution traces of one (or all) thread interleavings.
		This thesis considers any memory accesses not protected by a known trusted {\em mutex API}
		to be potential races;
		this may include atomic instructions such as {\tt xchg},
		which most all-data-races-are-bugs research does not include,
		but are necessary for {\em Iterative Deepening}'s {\em verification} guarantee
		(\sect{\ref{sec:quicksand-soundness}}).
	\item {\bf Deterministic bug.}
		As {\em concurrency bug}, but manifesting in all (or at least most) thread interleavings,
		making it easy for programmers to find and debug without advanced tooling.
		This thesis will occasionally use ``manifests on the 1st interleaving Landslide tested''
		as an approximation, without necessarily checking for its presence or absence in other branches.
	\item {\bf DPOR.} See {\em dynamic partial order reduction}.
	\item {\bf Dynamic partial order reduction}.
		Exploration algorithm for {\em stateless model checkers} which
		identifies alternate interleavings to explore likely to result
		in reorderings of {\em memory conflicts} observed in the current branch.
		Runs iteratively upon completion of each branch, tagging one or more sibling branches
		for future exploration, skipping over ones guaranteed to be {\em independent}.
		See \sect{\ref{sec:landslide-dpor}}.
	\item {\bf Equivalent.}
		Said of two {\em interleavings} where one can be transformed into the other using
		permutations of only {\em independent interleavings}.
		Note the difference in ``type signature'' with {\em independent}, which applies to {\em transitions} instead.
	\item {\bf Estimation.}
		An algorithm whose input is the known structure of a partially-explored {\em state space}
		and whose output is a guess at the ultimate size of the state space when exploration finish. % subjunctive woo
		Output may be in units of either time or branches.
		See \sect{\ref{sec:landslide-estimate}}.
	\item {\bf ETA.}
		Short for {\em estimation} (or ``state space estimate''),
		typically with nuance for units of time rather than branches.
	\item {\bf Execution tree.}
		% TODO
	\item {\bf False negative.}
		Said of a tool's failure to produce a bug report when a bug does exist after all.
		In theory this may mean that the {\em test case} might have been insufficient to expose the bug
		(i.e., the code under test is wrong in any way whatsoever);
		however, for pragmatic, tool-oriented discussion,
		this thesis will use the term to mean bugs observable under the given test case which the tool itself fails to find.
	\item {\bf False positive.}
		Said of a bug report which,
		upon subjective manual inspection by the programmer,
		does not indicate a bug after all.
		In principle, Landslide's bug-detection predicates always indicate something truly wrong,
		so it cannot produce false positive bug reports;
		in practice, however, infinite loop detection and deadlock detection in the presence of ad-hoc synchronization
		are both heuristic (\sect{\ref{sec:landslide-blocking}}),
		so may produce false positives after all, although I have never witnessed one to date.
	\item {\bf Happens-before.}
		% TODO
	\item {\bf Heuristic.}
		% TODO
	\item {\bf Independent.}
		% TODO
		Note the difference in ``type signature'' with {\em equivalent}, which applies to {\em interleavings} instead.
	\item {\bf Interleaving.}
		% TODO
	\item {\bf Iterative Deepening.}
		The heuristic search ordering strategy used by {\em Quicksand}.
		One of this thesis's major contributions.
		See \sect{\ref{sec:quicksand-id}}.
	\item {\bf Kernelspace.}
		Opposite of {\em userspace}.
		Execution mode in which code has the ability to modify critical system state.
		In addition to having access to privileged CPU instructions,
		kernelspace code is also responsible for implementing safe
		concurrent scheduling, safe virtual memory protection and isolation guarantees, et cetera.
	\item {\bf Landslide.}
		Your friendly local {\em stateless model checking} implementation, written by yours truly.
		The framework upon which all this thesis's contributions are built.
		Documented in Chapter \ref{chap:landslide}.
	\item {\bf Landslide-friendly.}
		Said of a {\em test case} that is well-suited for {\em model checking},
		i.e.,
		contains few or no loops which,
		under normal circumstances, would increase the probability of a bug manifesting,
		but in principle do not make any new behaviours possible that would be impossible without the loop.
	\item {\bf Limited happens-before.}
		% TODO
	\item {\bf Lock.}
		% TODO
	\item {\bf Memory conflict.} See {\em conflict}.
	\item {\bf mutex.}
		Short for {\em mutual exclusion lock}. The name {\em P2} and {\em pthreads} use for their {\em lock API}.
	\item {\bf Model checking.}
		In general, refers to the whole research field of checking program implementations against
		formal specifications in a theoretically rigorous way.
		In this thesis, used as shorthand for {\em stateless model checking}.
	\item {\bf P2.}
		% TODO
		{\em API} is very similar to that of {\em pthreads}.
		Often used here to refer to the same thread library project at both CMU and PSU,
		even though the latter did not call it by that name.
	\item {\bf pthreads.}
		Thread library for Linux programs provided by POSIX.
	\item {\bf Pure happens-before.}
		% TODO
	\item {\bf Quicksand.}
		% TODO
	\item {\bf Race condition.}
		A confusing term which, while CMU 15-410 teaches it to mean {\em concurrency bug},
		in research circles is generally taken to mean {\em data race}.
		This thesis will avoid it entirely, using those two more precise terms instead.
	\item {\bf Schedule.}
		Another term for {\em interleaving} with a more theoretical nuance.
	\item {\bf Stateless model checking.}
		% TODO
	\item {\bf State space.}
		% TODO
	\item {\bf Systematic testing.}
		Another name for {\em stateless model checking}.
	\item {\bf TA.} See {\em teaching assistant}.
	\item {\bf Teaching assistant.}
		% TODO
	\item {\bf Test case.}
		% TODO
	\item {\bf Thread.}
		% TODO
	\item {\bf Transition.}
		% not the kind of transition that's for people
		A sequence of execution steps by a single thread,
		characterized by the two different program states at beginning and end.
		In this thesis, will be used exclusively when the terminal program states are {\em preemption points},
		and no intermediate state is a preemption point as well.
	\item {\bf Userspace.}
		Opposite of {\em kernelspace}.
		Execution mode in which programs, such as {\em P2} or client code thereof,
		run with limited privileges,
		such as no access to the CPU's control registers, interrupt descriptor table, ability to perform device I/O,
		et cetera.
		Assumed to run with the virtual memory protections standard in any modern kernel
		(such as crashing when dereferencing a null pointer).
	\item {\bf Vector clock.}
		% TODO
	\item {\bf Verification.}
		Formally certifying the correctness of a program under certain properties.
		In this thesis, will always be constrained by the set of behaviours
		the {\em test case} is capable of exposing in the first place.
\end{enumerate}
